### 2020-05-01 final_report/data.tex
more minor edits

The expensive MC TBR model is fundamentally a Monte Carlo simulation based on the
OpenMC framework~\cite{ROMANO201590}. As input the software expects 18~parameters, discrete and
continuous, that are fully listed in~\cref{tbl:params}. During evaluation, which
usually takes units of seconds, a fixed number of events is simulated, and the
results are given in terms of the mean and the standard deviation of the~TBR aggregated
over the simulated run. The former of these two we accept to be the output TBR
value that is subject to approximation.

@@

In our work, we deliberately make no assumptions about the internal properties of the
MC TBR simulation, effectively treating it as a black box model. This limits our
means of studying its behaviour to inspection of its outputs at various
points in the feature space. We therefore require
sufficiently large and representative quantities of samples to ensure that
surrogates can be trained to approximate the MC TBR model.

With a grid search in such a high-dimensional domain clearly intractable, we
selected uniform pseudo-random sampling\footnote{Continuous and discrete
parameters are drawn from a corresponding uniform distribution over their
domain, as defined in~\cref{tbl:params}. For reproducibility, each run uses a seed equal to its number.} to generate large amounts of feature
configurations that we consider to be independent and unbiased. For evaluation
of the expensive MC TBR model, we utilise parallelisation offered by
the HPC infrastructure available at UCL computing facilities. To this end, we
designed and implemented the Approximate TBR Evaluator---a Python software package capable of
sequential evaluation of the multi-threaded OpenMC simulation on batches of
previously generated points in the feature space.
Having deployed ATE at the UCL Hypatia cluster,\footnote{The Hypatia RCIF
	partition is comprised of 4~homogeneous nodes. Each node is installed with
	376~GB RAM and 40~Intel\textsuperscript{\textregistered}
	Xeon\textsuperscript{\textregistered} Gold 6148 CPUs of clock frequency
	2.40~GHz.} we completed 3~data
generation runs that are summarised in~\cref{tbl:sampling-runs}.

\begin{table}[h]
	\sisetup{round-mode=places,round-precision=2}
	\centering
	{\footnotesize
		\begin{tabular}{rlllll}
		\toprule
		\#	& Samples & Batch division & $t_{\text{run}} $ 
			& $\overline{t}_{\text{eval.}}$ [\si{\second}] & Description \\
		\midrule
		0 & \num{100000} & $\num{100}\times\num{1000}$ & 2~days, 23~h 
		  & $\num{7.877122043981347} \pm \num{2.749485948560108}$ &
		Testing run using old MC TBR version.\\
		1 & \num{500000} & $\num{500}\times\num{1000}$ & 13~days, 20~h
		  & $\num{7.777049573054314} \pm \num{2.8103592103930337}$ &
		Fully uniform sampling in the entire domain.\\
		2 & \num{400000} & $\num{400}\times\num{1000}$ & 10~days 
		  & $\num{7.944379778103232} \pm \num{2.601428571503671}$ &
		Mixed sampling, discrete features fixed.\\
		\bottomrule
		\end{tabular}
	}
	\caption{Parameters of sampling runs. Here, $t_{\text{run}}$ denotes the total run
		time (including waiting in the processing queue), and
		$\overline{t}_{\text{eval.}}$ is the mean evaluation time of MC TBR (per single
		sampled point).}
	\label{tbl:sampling-runs}
\end{table}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{run1_500k_tbr_vs_feat}
	\caption{Marginalised dependence of TBR on the choice of continuous
	features in \num{500000}~points collected in run~1. Points are coloured by density.}
  \label{fig:tbr-vs-features}
\end{figure}

@@

\begin{wrapfigure}[15]{l}{0.25\textwidth}
	\centering
	{\footnotesize \incfigscale{0.8}{autoencoder}}
	\caption{Autoencoder with input width~$W_0$ and bottleneck width~$W_1$.
	Here, $\text{Dense}(N)$ denotes a fully-connected layer of $N$ neurons.}
	\label{fig:autoencoder}
\end{wrapfigure}

In particular, we consider two equally-sized sets\footnote{Each set contained
\num{100000}~samples from batches 0-99 of the corresponding runs.} of samples:
(a)~a subset of data obtained
from run~1 and (B)~A SUBSet of a single slice obtained from run~2. Our
expectation was that while the former case would provide meaningful insights into
correlations within the feature space, the latter would validate our
autoencoder implementation by analysing a set of points that are trivially
reducible in dimensionality due to a number of fixed discrete features.

@@

Confirming our expectation that in the latter, trivial case the
autoencoder should achieve greater dimensional reduction, we are inclined to
believe that our implementation is indeed operating as intended. However, we
must also conclude that in both examined cases this method failed to produce a
reduction that would be superior to a naïve approach.\footnote{In both tested cases we
	can trivially eliminate 7~dimensions due to overdetermined one-hot-encoded
	features and 2~dimensions due to sum-to-one constraints. Furthermore, in the
	single slice case we may omit 7~additional dimensions due to artificially fixed
features.} This is consistent with previous results obtained by PCA and kriging.

>>>>

The expensive MC TBR model is fundamentally a Monte Carlo simulation based on the
OpenMC framework~\cite{ROMANO201590}. As input the software expects 18~parameters, discrete and
continuous, that are fully listed in~\cref{tbl:params}. During evaluation, which
usually takes units of seconds, a fixed number of neutron events is generated, and the
results are given in terms of the mean and the standard deviation of the~TBR aggregated
over the simulated run. The former of these two we accept to be the output TBR
value that is subject to approximation.

@@

In our work, we deliberately make no assumptions about the internal properties of the
MC TBR simulation, effectively treating it as a black box model. This limits our
means of studying its behaviour to inspection of its outputs at various
points in the feature space. We therefore require
sufficiently large and representative quantities of samples to ensure that
surrogates can be trained to approximate the MC TBR model accurately.

With a grid search in such a high-dimensional domain clearly intractable, we
selected uniform pseudo-random sampling\footnote{Continuous and discrete
parameters are drawn from a corresponding uniform distribution over their
domain, as defined in~\cref{tbl:params}. For reproducibility, each run uses a seed equal to its number.} to generate large amounts of feature
configurations that we consider to be independent and unbiased. For evaluation
of the expensive MC TBR model, we utilise parallelisation offered by
the HPC infrastructure available at UCL computing facilities. To this end, we
designed and implemented the Approximate TBR Evaluator---a Python software package capable of
sequential evaluation of the multi-threaded OpenMC simulation on batches of
previously generated points in the feature space.
Having deployed ATE at the UCL Hypatia cluster\footnote{The Hypatia RCIF
	partition is comprised of 4~homogeneous nodes. Each node is installed with
	376~GB RAM and 40~Intel\textsuperscript{\textregistered}
	Xeon\textsuperscript{\textregistered} Gold 6148 CPUs of clock frequency
	2.40~GHz.}, we completed three data
generation runs that are summarised in~\cref{tbl:sampling-runs}.

\begin{table}[h]
	\sisetup{round-mode=places,round-precision=2}
	\centering
	{\footnotesize
		\begin{tabular}{rlllll}
		\toprule
		\#	& Samples & Batch division & $t_{\text{run}} $ 
			& $\overline{t}_{\text{eval.}}$ [\si{\second}] & Description \\
		\midrule
		0 & \num{100000} & $\num{100}\times\num{1000}$ & 2~days, 23~h 
		  & $\num{7.877122043981347} \pm \num{2.749485948560108}$ &
		Testing run using old MC TBR version.\\
		1 & \num{500000} & $\num{500}\times\num{1000}$ & 13~days, 20~h
		  & $\num{7.777049573054314} \pm \num{2.8103592103930337}$ &
		Fully uniform sampling in the entire domain.\\
		2 & \num{400000} & $\num{400}\times\num{1000}$ & 10~days 
		  & $\num{7.944379778103232} \pm \num{2.601428571503671}$ &
		Mixed sampling, discrete features fixed.\\
		\bottomrule
		\end{tabular}
	}
	\caption{Parameters of sampling runs. Here, $t_{\text{run}}$ denotes the total run
		time (including waiting in the processing queue), and
		$\overline{t}_{\text{eval.}}$ is the mean evaluation time of the MC TBR
		model (per single sampled point).}
	\label{tbl:sampling-runs}
\end{table}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{run1_500k_tbr_vs_feat}
	\caption{Marginalised dependence of TBR on the choice of continuous
	features in \num{500000}~points generated in run~1. Points are coloured by density.}
  \label{fig:tbr-vs-features}
\end{figure}

@@

\begin{wrapfigure}[15]{l}{0.25\textwidth}
	\centering
	{\footnotesize \incfigscale{0.8}{autoencoder}}
	\caption{Autoencoder with input width~$W_0$ and bottleneck width~$W_1$.
	Here, $\text{Dense}(N)$ denotes a fully-connected layer of $N$~neurons.}
	\label{fig:autoencoder}
\end{wrapfigure}

In particular, we consider two equally-sized sets\footnote{Each set contained
\num{100000}~samples from batches 0-99 of the corresponding runs.} of samples:
(a)~a~subset of data obtained
from run~1 and (b)~a~subset of a single slice obtained from run~2. Our
expectation was that while the former case would provide meaningful insights into
correlations within the feature space, the latter would validate our
autoencoder implementation by analysing a set of points that are trivially
reducible in dimensionality due to a number of fixed discrete features.

@@

Confirming our expectation that in the latter, trivial case the
autoencoder should achieve greater dimensional reduction, we are inclined to
believe that our implementation is indeed operating as intended. However, we
must also conclude that in both examined cases this method failed to produce a
reduction that would prove superior to a naïve approach.\footnote{In both tested cases we
	can trivially eliminate 6~dimensions due to overdetermination of one-hot-encoded
	categorical features, and 2~dimensions corresponding to sum-to-one constraints. Furthermore, in the
	single slice case we may omit 6~additional dimensions due to fixed feature
	assignment.} This is consistent with previous results obtained by PCA and kriging.

### 2020-05-01 final_report/introduction.tex
more minor edits

In this internship project, in coordination with the UK Atomic Energy Authority (UKAEA), we sought to develop a surrogate model for the tritium breeding ratio (TBR) in a Tokamak nuclear fusion reactor. Our expensive model was an MC-based neutronics simulation, \textit{Paramak}\footnote{Provided by collaborator Jonathan Shimwell, at UKAEA.}, which returns a prediction of the TBR for a given configuration of a spherical Tokamak. We took an empirical approach to the construction of this surrogate, and no results described here are explicitly dependent on prior physics knowledge.

@@

The input parameters of the computationally-expensive TBR model therefore fall
into two classes. Continuous parameters, including material thicknesses and
packing ratios, describe the geometry of a given reactor configuration. Discrete
categorical parameters further specify all relevant material sections, including
coolants, armours, and neutron multipliers. One notable exception is the
enrichment ratio, a continuous parameter denoting the presence of
\isotope[6]{Li}. Our challenge, put simply, was to produce a TBR function which
takes these same input parameters and approximates the MC TBR model with the
greatest achievable regression performance.

>>>>

In this project, in coordination with the UK Atomic Energy Authority (UKAEA), we sought to develop a surrogate model for the tritium breeding ratio (TBR) in a Tokamak nuclear fusion reactor. Our expensive model was an MC-based neutronics simulation, \textit{Paramak}\footnote{Provided by collaborator Jonathan Shimwell, at UKAEA.}, which returns a prediction of the TBR for a given configuration of a spherical Tokamak. We took an empirical approach to the construction of this surrogate, and no results described here are explicitly dependent on prior physics knowledge.

@@

The input parameters of the computationally-expensive TBR model therefore fall
into two classes. Continuous parameters, including material thicknesses and
packing ratios, describe the geometry of a given reactor configuration. Discrete
categorical parameters further specify all relevant material sections, including
coolants, armours, and neutron multipliers. One notable exception is the
enrichment ratio, a continuous parameter denoting the presence
of~\isotope[6]{Li}. Our challenge, put simply, was to produce a fast TBR
function which takes these same input parameters and approximates the MC TBR
model with the greatest achievable regression performance.

### 2020-05-01 final_report/methodology.tex
more minor edits

In the decoupled sampling approach that is further described
in~\cref{sec:supervised}, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on specific model family and appropriate choice of its
hyperparameters, surrogate models $\hat{f}(x)$ are trained to minimise
the empirical risk~$R_{\text{emp.}}$ with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where the empirical risk is defined as
$R_{\text{emp.}}(\hat{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\hat{f}(x^{(i)}),f(x^{(i)})\right)$.

The adaptive sampling approach that we characterise in~\cref{sec:adaptive} can be viewed as a more general problem.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, multiple sets $\{\mathcal{T}_k\}_{k=0}^K$ are used, such that
$\mathcal{T}_{k-1}\subset\mathcal{T}_k$ for all $k>1$. The first set
$\mathcal{T}_0$ is initialised randomly to provide a \textit{burn-in}, and is
repeatedly extended in epochs, whereby each epoch trains a new surrogate on
$\mathcal{T}_k$ using the supervised learning procedure, evaluates its
performance, and forms a new set $\mathcal{T}_{k+1}$ by adding more samples to
$\mathcal{T}_k$. This permits the learning algorithm to condition the selection
of new samples on the evaluation results in order to maximise improvement of
surrogate performance over complex regions within the domain.

@@

\Cref{tbl:metrics} provides an exhaustive listing and description of metrics recorded
in the experiments. For regression performance analysis, we include a selection
of absolute metrics to assess the approximation capability of surrogates, and set
practical bounds on the expected uncertainty of their predictions. In addition, we also track
relative measures that are better-suited for comparison between this work and others, as
they maintain invariance with respect to the selected domain and image space.
For complexity analysis, surrogates are assessed in terms of wall time
elapsed during training and predicition. This is motivated by common practical use
cases of our work, where models are trained and used as drop-in replacements for the
expensive MC TBR model. Since training set sizes remain to be determined, all times are
reported per a single datapoint. Even though some surrogates support acceleration
by means of parallelisation, measures were taken to ensure sequential
processing of samples to achieve comparability between considered
models.\footnote{The only exception to this are artificial neural networks, which are
notoriously slow to train on conventional CPU architectures in serialised
setting.}

@@

To prevent undesirable bias in results due to training set selection, all metrics
are collected in the scheme of $k$-fold cross-validation with a conventional choice of
$k=5$. In this setting, a sample set is divided into 5 disjoint folds, each of which
is used as a test set for the model trained on the remaining 4.\footnote{Unless explicitly stated otherwise, we use 1
fold for testing and the 4 remaining folds for training. This gives 80\% to
20\% train-test ratio.} In each such run experiments are
repeated, and the overall value of each metric is reported as the mean and
standard deviation over all folds.

@@

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in~\cref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate families to draw from. This is described by the listing in~\cref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomised algorithms, artificial neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these families, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

While some of the presented model families are clearly determined by an explicit
choice of a learning algorithm, others may vary considerably depending on
hyperparameter configuration. A good example of the latter group are artificial neural
networks, which in addition to conventional hyperparameters enable to control network
architecture through selection from various parametric graph templates
(see~\cref{fig:nn-archs} for details). This allows us to realise a simplistic
network architecture search during the hyperparameter optimisation procedure.

@@

The goal of the last two experiments is to exploit the information gathered by
hyperparameter tuning to investigate scaling properties. In
the third experiment, the 20~best-performing hyperparameter choices of each family
(in $R^2$) are used to train surrogates using progressively larger training sets. Following
that, the fourth experiment aims to produce surrogates for practical use by
retraining selected well-scaling models to further optimise their properties.

>>>>

In the decoupled sampling approach that is further described
in~\cref{sec:supervised}, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on a specific model family and appropriate choice of its
hyperparameters, surrogate models $\hat{f}(x)$ are trained to minimise
the empirical risk~$R_{\text{emp.}}$ with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where the empirical risk is defined as
$R_{\text{emp.}}(\hat{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\hat{f}(x^{(i)}),f(x^{(i)})\right)$.

The adaptive sampling approach that we characterise in~\cref{sec:adaptive} can be viewed as a more general problem.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, multiple sets $\{\mathcal{T}_k\}_{k=0}^K$ are used, such that
$\mathcal{T}_{k-1}\subset\mathcal{T}_k$ for all $k>1$. The first set
$\mathcal{T}_0$ is initialised randomly to provide a \textit{burn-in}, and is
repeatedly extended in epochs, whereby each epoch trains a new surrogate~$\hat{f}_k(x)$ on
$\mathcal{T}_k$ using the supervised learning procedure, evaluates its
performance, and forms a new set $\mathcal{T}_{k+1}$ by adding more samples to
$\mathcal{T}_k$. This permits the learning algorithm to condition the selection
of new samples in~$\mathcal{T}_{k+1}$ on the evaluation results
of~$\hat{f}_k(x)$ in order to maximise improvement of
surrogate performance over complex regions within the feature space.

@@

\Cref{tbl:metrics} provides an exhaustive listing and description of metrics recorded
in our experiments. For regression performance analysis, we include a selection
of absolute metrics to assess the approximation capability of surrogates, and set
practical bounds on the expected uncertainty of their predictions. In addition, we also track
relative measures that are better-suited for comparison between this work and others, as
they maintain invariance with respect to the selected domain and image space.
For complexity analysis, surrogates are assessed in terms of wall
time\footnote{Real time elapsed during computation measured by a
chronometer, here by means of the Python~\texttt{time} package.}
elapsed during training and predicition. This is motivated by common practical use
cases of our work, where models are trained and used as drop-in replacements for the
expensive MC TBR model. Since training set sizes remain to be determined, all times are
reported per a single datapoint. Even though some surrogates support acceleration
by means of parallelisation, sequential processing of samples was ensured to
achieve comparability between considered
models.\footnote{The only exception to this are artificial neural networks,
	which require considerable amount of processing power for training on conventional CPU architectures.}

@@

To prevent undesirable bias in results due to training set selection, all metrics
are collected in the scheme of $k$-fold cross-validation with a conventional choice of
$k=5$. In this setting, a sample set is uniformly divided into 5 disjoint folds, each of which
is used as a test set for models trained on the remaining 4.\footnote{Unless explicitly stated otherwise, we use 1
fold for testing and the 4 remaining folds for training. This gives 80\% to
20\% train-test ratio.} Having repeated the same experiment for each such run,
the overall value of individual metrics is reported in terms of their mean and
standard deviation over all folds.

@@

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in~\cref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate families to review. This is described by the listing in~\cref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomised algorithms, artificial neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these families, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

While some of the presented model families are clearly determined by an explicit
choice of a learning algorithm, others may vary considerably depending on
hyperparameter configuration. A good example of the latter group are artificial neural
networks, which in addition to conventional hyperparameters enable to control network
architecture through selection from various parametric graph templates
(illustrated in~\cref{fig:nn-archs}). This allows us to realise a simplistic
network architecture search during the hyperparameter tuning procedure.

@@

The objective of the last two experiments is to exploit the information gathered by
hyperparameter tuning. In the third experiment, the 20~best-performing
hyperparameter configurations of each family (with respect to~$R^2$) are used to
perform training on progressively larger sets to investigate their scaling
properties. Following that, the fourth experiment aims to produce surrogates
suitable for practical use by retraining selected well-scaling instances on large
training sets to satisfy the goals of this work.

### 2020-05-01 final_report/results.tex
more minor edits

We begin by comparing a diverse set of surrogate families that we proposed
to evaluate on previously collected samples of the expensive MC TBR model.
Through the four experimental cases described
in~\cref{sec:experiment-methodology}, we aim to study properties of the
considered models in terms of regression performance, training and prediction
time.

@@

The first two experiments perform Bayesian optimisation to maximise~$R^2$ in
a cross-validation setting as a function of model hyperparameters. While in the
first experiment we limit training and test sets to the scope of four selected
slices of the feature space, in the second experiment we lift this restriction.

@@

The results of the second experiment, shown in~\cref{fig:exp2-time-vs-reg} (and listed
in~\cref{tbl:exp2-detailed-results} in the Appendix),
seem to confirm our expectations. Compared to the previous case, we observe
that many surrogate families consistently achieved worse regression
performance and prediction times in a more complex, unrestricted domain. The least
affected models appear to be GBTs, ANNs and ERTs, which are known to be capable of capturing relationships
involving mixed feature types that were deliberately withheld in the first
experiment. With only negligible differences, the first two of these families
appear to be tied for the best performance as well as the shortest prediction
time. We observe that ERTs trees and RBFs also
demonstrated satisfactory results, outperforming the remaining surrogates in
terms of regression performance, and in some cases also in prediction time.

\begin{wrapfigure}[12]{r}{0.333\textwidth}
	\centering
	\vspace{-2ex}
	\includegraphics[width=\linewidth]{exp2_time_vs_reg}
	\caption{Results of experiment~2, plotted analogously
	to~\cref{fig:exp1-time-vs-reg}.}
	\label{fig:exp2-time-vs-reg}
\end{wrapfigure}

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. Firstly, the results shown 
in~\cref{fig:scaling-r2} (and listed in~\cref{tbl:exp3-detailed-results-r2} in
the Appendix) suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training data. While such families achieve nearly comparable regression
performance on the largest dataset, in the opposite case methods based on
decision trees clearly outperform neural networks. This can be observed
particularly on training sets of sizes up to~\num{6000}.

@@

Next, we examine scaling behaviour in terms of the mean training time (displayed
in~\cref{fig:scaling-trn} and listed in~\cref{tbl:exp3-detailed-results-t-train} in
the Appendix). Consistent with our expectation, the shortest times
were achieved by instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also exhibit
desirable properties unlike RBFs and GPR, which appear to scale superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

Finally, we study scaling with respect to the mean prediction time (shown
in~\cref{fig:scaling-pred} and listed in~\cref{tbl:exp3-detailed-results-t-pred} in
the Appendix). Our initial observation is that all tested
surrogate families with the exception of previously mentioned instance-based
learning methods offer desirable characteristics overall. Analogous to previous
experiments, GBTs and ANNs again appear to be tied, as they not only exhibit comparable times but also similar
scaling slopes.

@@

Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameter
assignments. Trained to satisfy
goal~(a), the best surrogate (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline measured during run~1
(see~\cref{tbl:sampling-runs} for details). Satisfying
goal~(b), the fastest model (no.~2, ANN) features~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, pursuing
goal~(c) we also trained a simplified model (no.~5, GBT) that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with only a set of size~\num{10000}.

>>>>

We begin by comparing a diverse set of surrogate families that we proposed
to evaluate on previously generated samples of the expensive MC TBR model.
Through the four experimental cases described
in~\cref{sec:experiment-methodology}, we aim to study properties of the
considered models in terms of regression performance, training and prediction
time.

@@

The first two experiments perform Bayesian optimisation to maximise~$R^2$ in
a cross-validation setting as a function of model hyperparameters. While in the
first experiment we limit training and test sets to the scope of four selected
slices of the feature space, in the second experiment we lift this restriction
to examine surrogate capability to model a more complex domain.

@@

The results of the second experiment, shown in~\cref{fig:exp2-time-vs-reg} (and listed
in~\cref{tbl:exp2-detailed-results} in the Appendix),
seem to confirm our expectations. Compared to the previous case, we observe
that many surrogate families consistently achieved worse regression
performance and prediction times. The least
affected models appear to be GBTs, ANNs and ERTs, which are known to be capable of capturing relationships
involving mixed feature types that were deliberately withheld in the first
experiment. With only negligible differences, the first two of these families
appear to be tied for the best performance as well as the shortest prediction
time. We observe that ERTs and RBFs also
demonstrated satisfactory results, relatively outperforming the remaining surrogates in
terms of regression performance, and in some cases also in prediction time.

\begin{wrapfigure}[10]{r}{0.333\textwidth}
	\centering
	\vspace{-2ex}
	\includegraphics[width=\linewidth]{exp2_time_vs_reg}
	\caption{Results of experiment~2, plotted analogously
	to~\cref{fig:exp1-time-vs-reg}.}
	\label{fig:exp2-time-vs-reg}
\end{wrapfigure}

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. Firstly, the results shown 
in~\cref{fig:scaling-r2} (and listed in~\cref{tbl:exp3-detailed-results-r2} in
the Appendix) suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable regression
performance on the largest dataset, in the opposite case methods based on
decision trees clearly outperform neural networks. This can be observed
particularly on sets of sizes up to~\num{6000}.

@@

Next, we examine scaling behaviour in terms of the mean training time (displayed
in~\cref{fig:scaling-trn} and listed in~\cref{tbl:exp3-detailed-results-t-train} in
the Appendix). Consistent with our expectation, the shortest times
were achieved by instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

Finally, we study scaling with respect to the mean prediction time (shown
in~\cref{fig:scaling-pred} and listed in~\cref{tbl:exp3-detailed-results-t-pred} in
the Appendix). Our initial observation is that all tested
surrogate families with the exception of previously mentioned instance-based
learning methods offer desirable characteristics overall. Analogous to previous
experiments, GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following that, we notice a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

@@

Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameter
assignments. In pursuit of goal~(a), the best approximator (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error of regression~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline measured during run~1
(see~\cref{tbl:sampling-runs} for details). Satisfying
goal~(b), the fastest model (no.~2, ANN) features~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, in our
attempts to satisfy goal~(c) we also trained a more simplified model (no.~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with only a set of size~\num{10000}.
